---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# nzbabynames

The nzbabynames package provides the following datasets

- `nzbabynames` This data set includes the sex and count of each first name from 1900 to 2017. The data set includes all names with more than 10 instances in any given year. Source: https://catalogue.data.govt.nz/dataset/baby-name-popularity-over-time/

```{r print-nz-babynames}
library(nzbabynames)
head(nzbabynames)
```

- `maorinames` This data set includes the  top 20 Māori baby names for 2016 and 2015 by sex. In 2015 data only includes children born in NZ, in 2016 rankings Māori baby names among children born overseas are also included Source: https://catalogue.data.govt.nz/dataset/most-popular-male-and-female-first-names

```{r print-maori-names}
head(maorinames)
```

- `nzbirths` Data on live births from 1935-2017. Data from 1971 includes live births by sex. *Note: Data before 1991 includes New Zealand residents and overseas visitors. Data from 1991 includes New Zealand residents only.* (Source: Stats NZ and licensed by Stats NZ for re-use under the Creative Commons Attribution 4.0 International licence.")

```{r print-nzbirths}
head(nzbirths)
```

## Installation

You can install the released version of nzbabynames from github:

```{r install, eval = FALSE}
install_github("ropenscilabs/nzbabynames")
```

