---
title: "How to Tell Someone's Age When All You Know Is Her Name"
author: "Emily Kothe"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    keep_md: yes
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette is based on this FiveThirtyEight article [How to Tell Someone’s Age When All You Know Is Her Name](https://fivethirtyeight.com/features/how-to-tell-someones-age-when-all-you-know-is-her-name/#fn-1). That article demonstrated how you could use actuarial life tables and a database of baby names to estimate the age of living Americans with a given name.

In this vignette we use the nzbabynames data set and the NZ complete cohort actuarial life tables to calculate the median age of individuals with a given first name.^[Like FiveThirtyEight we assume that people of the same sex die at the same rate regardless of their name - in practice will vary by a number of factors related to both name and life expectancy such as ethnicity].

## An "old" name

Let's start with one name that is sterotypically 'old' - Ethel

```{r message=FALSE, warning = FALSE}
library(ggplot2)
library(matrixStats)
library(dplyr)
library(nzbabynames)
library(stringr)
library(lubridate)

name_of_interest <- "Ethel"

df <- life_tables %>% 
  filter(percentile == "median") %>% 
  filter(age == (lubridate::year(Sys.Date()) - yearofbirth)) %>% 
  mutate(prop_still_alive = lx/100000, 
         Year = yearofbirth, 
         Sex = stringr::str_to_title(sex)) %>% 
  right_join(filter(nzbabynames, Name == name_of_interest), by= c("Year", "Sex")) %>% 
  mutate(prop_still_alive = case_when(
    Year < 1918 ~ 0, 
    TRUE ~ prop_still_alive), 
  Name = case_when(
    is.na(Name) ~ name_of_interest, 
    TRUE ~ Name), 
  Count = case_when(
    is.na(Count) ~ 0, 
    TRUE ~ as.numeric(Count)),
  alive_count = Count * prop_still_alive)

med_age <- df %>% 
  summarise(wted_median = weightedMedian(age, w = alive_count, na.rm = TRUE))

max <- max(df$Count)

peak <- df %>% 
  filter(Count == max)

ggplot(data = df, aes(x = Year)) +
  geom_col(aes(y = Count), fill="grey")+
  geom_col(aes(y = alive_count), fill="cornflowerblue") +
  scale_x_continuous(limits = c(1900, 2016)) +
  coord_cartesian(expand = FALSE) +
  theme_classic()
```

`r name_of_interest` peaked in `r peak$Year` when `r max` `r name_of_interest`s were born. Based on the actuarial data, `r round(sum(df$alive_count)/sum(df$Count)*100,0)`% of `r name_of_interest`s born between 1900 and 2017 would still be alive in `r lubridate::year(Sys.Date())`. The median age of an `r name_of_interest` still alive, would be `r round(med_age$wted_median, 0)` years.

## And a young one

Let's contrast that with a 'younger' name - Brittany. 

```{r message=FALSE, warning = FALSE}
name_of_interest <- "Brittany"

df <- life_tables %>% 
  filter(percentile == "median") %>% 
  filter(age == (lubridate::year(Sys.Date()) - yearofbirth)) %>% 
  mutate(prop_still_alive = lx/100000, 
         Year = yearofbirth, 
         Sex = stringr::str_to_title(sex)) %>% 
  right_join(filter(nzbabynames, Name == name_of_interest), by= c("Year", "Sex")) %>% 
  mutate(prop_still_alive = case_when(
    Year < 1918 ~ 0, 
    TRUE ~ prop_still_alive), 
  Name = case_when(
    is.na(Name) ~ name_of_interest, 
    TRUE ~ Name), 
  Count = case_when(
    is.na(Count) ~ 0, 
    TRUE ~ as.numeric(Count)),
  alive_count = Count * prop_still_alive)

med_age <- df %>% 
  summarise(wted_median = weightedMedian(age, w = alive_count, na.rm = TRUE))

max <- max(df$Count)

peak <- df %>% 
  filter(Count == max)

ggplot(data = df, aes(x = Year)) +
  geom_col(aes(y = Count), fill="grey")+
  geom_col(aes(y = alive_count), fill="cornflowerblue") +
  scale_x_continuous(limits = c(1900, 2016)) +
  coord_cartesian(expand = FALSE) +
  theme_classic()
```

`r name_of_interest` peaked in `r peak$Year` when `r max` `r name_of_interest`s were born. Based on the actuarial data, `r round(sum(df$alive_count)/sum(df$Count)*100,0)`% of `r name_of_interest`s born between 1900 and 2017 would still be alive in `r lubridate::year(Sys.Date())`. The median age of an `r name_of_interest` still alive, would be `r round(med_age$wted_median, 0)` years.

## Finally a less informative name

Ethel and Brittany are both names with pretty clear trajectories. But of course some names rise and fall over time. Let's pick one of these - Joseph 

```{r message=FALSE, warning = FALSE}
name_of_interest <- "Joseph"

df <- life_tables %>% 
  filter(percentile == "median") %>% 
  filter(age == (lubridate::year(Sys.Date()) - yearofbirth)) %>% 
  mutate(prop_still_alive = lx/100000, 
         Year = yearofbirth, 
         Sex = stringr::str_to_title(sex)) %>% 
  right_join(filter(nzbabynames, Name == name_of_interest), by= c("Year", "Sex")) %>% 
  mutate(prop_still_alive = case_when(
    Year < 1918 ~ 0, 
    TRUE ~ prop_still_alive), 
  Name = case_when(
    is.na(Name) ~ name_of_interest, 
    TRUE ~ Name), 
  Count = case_when(
    is.na(Count) ~ 0, 
    TRUE ~ as.numeric(Count)),
  alive_count = Count * prop_still_alive)

med_age <- df %>% 
  summarise(wted_median = weightedMedian(age, w = alive_count, na.rm = TRUE))

max <- max(df$Count)

peak <- df %>% 
  filter(Count == max)

ggplot(data = df, aes(x = Year)) +
  geom_col(aes(y = Count), fill="grey")+
  geom_col(aes(y = alive_count), fill="cornflowerblue") +
  scale_x_continuous(limits = c(1900, 2016)) +
  coord_cartesian(expand = FALSE) +
  theme_classic()
```

`r name_of_interest` peaked in `r peak$Year` when `r max` `r name_of_interest`s were born. Based on the actuarial data, `r round(sum(df$alive_count, na.rm=TRUE)/sum(df$Count, na.rm=TRUE)*100,0)`% of `r name_of_interest`s born between 1900 and 2017 would still be alive in `r lubridate::year(Sys.Date())`. The median age of an `r name_of_interest` still alive, would be `r round(med_age$wted_median, 0)` years.

However, note that while there is a clear peak in `r peak$Year` this name has had much more consistent consistent staying power than either Ethel or Brittany, meaning that you might be less confident about guessing the age of any given `r name_of_interest` based on name alone. Put more formally, the interquartile range (and thus uncertainty) around the estimated age for `r name_of_interest` is much greater than around Ethel or Brittany.
