---
title: "Top 10 Names"
author: "Emily Kothe"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Top 10 Names}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

One popular way to understand baby names is to look at the most popular names 
in any given year. 

Here let's look at the Top 10 names for each year in our data set.

Start by reading in the data and creating a rank for each name by Sex and Year.

```{r}
library(nzbabynames)
library(dplyr)


names <- nzbabynames

names <- names %>% 
  group_by(Sex, Year) %>% 
  arrange(-Count) %>% 
  mutate(Rank = seq_along(Count))

names

```

Now let's filter the data set so we only have names in the Top 10 in any given year

```{r}
names_top <- names %>%
  filter(Rank <= 10) 

tail(names_top)
```

We can find the most popular names in a given year

```{r}
names_top %>% 
  filter(Year == 1985, Rank == 1)
```

I could also plot names over time. Here I'll use ggplot2 and gganimate to plot 
male and female names separately so it's easier to read.
```{r}
library(ggplot2)
library(gganimate)

names_top_female <- names_top %>%
  filter(Sex == "Female") 

time <- ggplot(names_top_female, aes(x = Rank, y = Count, fill = Name)) +
  geom_col() + 
  facet_wrap("Sex") +
  geom_text(aes(label = Name), vjust = -0.5, size = 3) +
  theme_classic() +
  theme(legend.position="none") +
  labs(title="{current_frame}") +
  transition_manual(Year)

animate(time, duration=100)
```

```{r}
library(ggplot2)
library(gganimate)

names_top_male <- names_top %>%
  filter(Sex == "Male") 

time <- ggplot(names_top_male, aes(x = Rank, y = Count, fill = Name)) +
  geom_col() + 
  facet_wrap("Sex") +
  geom_text(aes(label = Name), vjust = -0.5, size = 3) +
  theme_classic() +
  theme(legend.position="none") +
  labs(title="{current_frame}") +
  transition_manual(Year)

animate(time, duration=100)
```

